### 2016-07-08 lpshort.tex
Chandru's shortened version

%!TEX root =  autocontgrlp.tex
\section{Approximate Linear Programming}
The LP formulation in \eqref{mdplp} can be represented in short as,
\begin{align}
\begin{split}
\label{mdplpshort}
\min_{J\in \R^n}\, &c^\top J\\
\text{s.t.}\mb &J\geq T J,
\end{split}
\end{align}
or 
%\text{ or }\nn\\
\begin{align}
\begin{split}
\label{shortalter}
\min_{J\in \R^n} \,&c^\top J\\
\text{s.t.}\mb &EJ\geq H J,
\end{split}
\end{align}
where $J\geq TJ$ is a shorthand for the $nd$ constraints in \eqref{mdplp} and $E$ is an $nd\times n$ matrix given by $E=[I,\ldots,I]^\top$, i.e., $E$ is obtained by stacking $d$ identical $n\times n$ identity matrices one over the other. Note that \eqref{mdplpshort} and \eqref{shortalter} are identical programs and differ only in notation. We use notation of type \eqref{mdplpshort} whenever we prefer brevity and we use notation \eqref{shortalter} in some definitions and proofs for the sake of clarity.
Recall that the vector $c\in \R^n$ must have positive entries to guarantee that the unique solution to the above LP is the optimal value function. By convention, we will also assume that $c$ is also normalized:
\begin{assumption}\label{probdist}
$c=(c(i),i=1,\ldots,n)\in \R^n$ is a positive probability distribution, i.e., $c(i)>0$ and $\sum_{i=1}^n c(i)=1$.
\todoc{I am pretty sure the positivity is needed.}
\end{assumption}

The approximate linear program (ALP) is obtained by making use of LFA in the LP, i.e., by adding the extra constraint $J=\Phi r$ in \eqref{mdplpshort} with $\Phi \in \R^{n\times k}$ and introducing 
the new variables $r\in \R^k$. 
By substitution, this leads to
\begin{align}\label{alp}
\begin{split}
\min_{r\in \R^k}\, &c^\top \Phi r\\
\text{s.t.}\mb &\Phi r\geq T \Phi r.
\end{split}
\end{align}
Unless specified otherwise we use $\tr$ to denote any solution 
to the ALP and $\tj=\Phi \tr$ to denote the corresponding approximate value function. 
The extra constraint added that $J$ lies in the (column) span of $\Phi$ could lead to an infeasible LP. An easy way to prevent this is if the vector $\one$ whose components are all equal to one is in the span of $\Phi$.
Indeed, by \cref{shift}, for any $t\in \R$, $T (t\one ) =  \alpha t \one+T \zero \le \alpha t \one + \max_{s,a} g_a(s)$, hence $t\one \ge \alpha t \one+T\zero $ happens as soon as
$t\ge \frac{1}{1-\alpha}\max_{s,a} g_a(s)$. 
This leads to the following assumption:
\begin{assumption}\label{one}
The first column of the feature matrix $\Phi$ (i.e., $\phi_1$) is $\one \in \R^n$. 
In other words, the constant function is part of the basis.
\end{assumption}
Under this assumption \eqref{alp} is thus feasible and since \eqref{alp} is obtained by adding an extra constraint to the LP \eqref{mdplpshort}, the solutions to  \eqref{alp} are bounded and $\tj$ is uniquely defined.

We will now recall two results due to de Farias and Van Roy 
	that bound the error due to the introduction of the extra constraint in the ALP.
For this, we will need the definition of weighted $L_1$ and $L_\infty$ norms.
For any $c,\rho:S \to \R$ positive valued functions
the weighted $L_1$-norms $\norm{\cdot}_{1,c}$ 
and 
the weighted $L_\infty$-norms  $\norm{\cdot}_{\infty,\rho}$ are defined as
\begin{align*}
||J||_{1,c}=\sum_{s \in S} c(s)|J(s)|\,, \qquad
||J||_{\infty,\rho}=\max_{s \in S} \frac{|J(s)|}{\rho(s)}\,, \qquad J:S\to \R\,.
\end{align*}

@@

%In Definition~\ref{modnorm}, the use of the weighting function $\rho$ allows us to measure the error taking into account the relative importance of the various states. A higher value of $\rho(s)$ means that the state $s$ is less important and vice-versa.\\
With this, we can recall the said result.  %below  which bounds the error in the approximate value function.
\begin{theorem}[Theorem~$4.2$ of \cite{ALP}]
\label{restateval}
Let $\tr$ be the solution to the ALP in \eqref{alp}, $\tilde{J}_c=\Phi \tilde{r}_c$.
% $\psi$ be as in \cref{lyap} 
%and $c$ be a distribution as in \eqref{probdist}. 
Then, it holds that
\begin{align*}
||J^*-\tj||_{1,c}\leq \frac{2c^\top \psi}{1-\beta_{\psi}}\min_{r}||J^*-\Phi r||_{\mn}\,.
\end{align*}
\end{theorem}
The next result  the loss in performance of a policy that is greedy w.r.t. $J_{\tu}$:
\begin{theorem}[Theorem~$3.1$ of \cite{ALP}]
\label{restatepol}
Let $\tilde{u}$ be the greedy policy with respect to the solution $\tj$ of the ALP. Then,
\begin{align*}
||J^*-J_{\tu}||_{1,c}\leq \frac{1}{1-\alpha}||J^*-\tj||_{1,c'}\,,
\end{align*}
where $c'=(1-\alpha)c^\top(I-\alpha P_{\tu})^{-1}$.
\end{theorem}
Theorems~\ref{restateval} and ~\ref{restatepol} together imply that the ALP addresses both the control and prediction problems. Please refer to \cite{ALP} for a more detailed treatment of the ALP.\\
Note that the ALP is a linear program in $k$ ($\ll n$) variables as opposed to the LP in \eqref{mdplpshort} which has $n$ variables. Nevertheless, the ALP has $nd$ constraints (same as the LP) which is an issue when $n$ is large and calls for constraint approximation/reduction techniques.
\todoc{Column generation, Dantzig-Wolf decomposition?}

>>>>

%!TEX root =  autocontgrlp.tex
\section{Approximate Linear Programming}
\begin{comment}
The approximate linear program (ALP) is obtained by making use of LFA in the LP, i.e., by adding the extra constraint $J=\Phi r$ in \eqref{mdplpshort} with $\Phi \in \R^{n\times k}$ and introducing 
the new variables $r\in \R^k$. 
By substitution, this leads to
\begin{align}\label{alp}
\begin{split}
\min_{r\in \R^k}\, &c^\top \Phi r\\
\text{s.t.}\mb &\Phi r\geq T \Phi r.
\end{split}
\end{align}
Unless specified otherwise we use $\tr$ to denote any solution 
to the ALP and $\tj=\Phi \tr$ to denote the corresponding approximate value function. 
\end{comment}
For any function $\chi\colon S\ra \R_+$, define the discounted maximal inflation of $\chi$ due to $P = (p_a)_{a\in A}$ as
\begin{align*}
\beta_{\chi}=\max_{s \in S} \frac{\underset{a \in A}{\max}\big(\alpha\sum_{s'}p_a(s,s')\chi(s')\big)}{\chi(s)}\,.
\end{align*}
\begin{definition}
A function $\chi:S\to\R_+$ is said to be a \emph{Lyapunov} function for $P = (p_a)_{a\in A}$ 
	if $\beta_{\chi}<1$.
\end{definition}
The error bound will be given in terms of a Lyapunov function $\psi$ for $P$ that is in the column span of $\Phi$:
\begin{assumption}\label{lyap}
$\psi\colon S \ra \R_+$ is a Lyapunov function for $P$
and is present in the column span of the feature matrix $\Phi$: For some $r_0\in \R^k$, $\Phi r_0 = \psi$.
\end{assumption}
It is straightforward to check that the vector $\one$ when viewed as an $S \to \R_+$ function 
is a Lyapunov function. Further, by \cref{one}, $\one$ is trivially present in the column span of $\Phi$, 
hence, \cref{lyap} is not limiting.
In what follows we will always assume that Assumptions \ref{probdist}--\ref{lyap} hold. 
When \cref{lyap} holds, it follows that for any $J\in \R^n$, $t>0$, $s\in S$,
\begin{align*}
(T(J+ t \psi))(s) &= 
\max_{a} g_a(s) + \alpha \sum_{s'} p_a(s,s') J(s') + t \alpha  \sum_{s'} p_a(s,s') \psi(s') \\
& \le 
\max_{a} g_a(s) + \alpha \sum_{s'} p_a(s,s') J(s') + t \beta_{\psi} \psi(s) \\
& = (T J)(s) + \beta_{\psi}\,t\,  \psi(s),
\end{align*}
or, in short,
\begin{align}
\label{eq:psilin}
T(J+ t \psi ) \le TJ + \beta_{\psi}\,t\,  \psi\,\, \quad (J\in \R^n,\, t>0)\,.
\end{align}
%Under this assumption \eqref{alp} is thus feasible and since \eqref{alp} is obtained by adding an extra constraint to the LP \eqref{mdplpshort}, the solutions to  \eqref{alp} are bounded and $\tj$ is uniquely defined.
\begin{comment}
We will now recall two results due to de Farias and Van Roy 
	that bound the error due to the introduction of the extra constraint in the ALP.
For this, we will need the definition of weighted $L_1$ and $L_\infty$ norms.
For any $c,\rho:S \to \R$ positive valued functions
the weighted $L_1$-norms $\norm{\cdot}_{1,c}$ 
and 
the weighted $L_\infty$-norms  $\norm{\cdot}_{\infty,\rho}$ are defined as
\begin{align*}
||J||_{1,c}=\sum_{s \in S} c(s)|J(s)|\,, \qquad
||J||_{\infty,\rho}=\max_{s \in S} \frac{|J(s)|}{\rho(s)}\,, \qquad J:S\to \R\,.
\end{align*}

@@

%In Definition~\ref{modnorm}, the use of the weighting function $\rho$ allows us to measure the error taking into account the relative importance of the various states. A higher value of $\rho(s)$ means that the state $s$ is less important and vice-versa.\\
With this, we can recall the said result.  %below  which bounds the error in the approximate value function.
\begin{theorem}[Theorem~$4.2$ of \cite{ALP}]
\label{restateval}
Let $\tr$ be the solution to the ALP in \eqref{alp}, $\tilde{J}_c=\Phi \tilde{r}_c$.
% $\psi$ be as in \cref{lyap} 
%and $c$ be a distribution as in \eqref{probdist}. 
Then, it holds that
\begin{align*}
||J^*-\tj||_{1,c}\leq \frac{2c^\top \psi}{1-\beta_{\psi}}\min_{r}||J^*-\Phi r||_{\mn}\,.
\end{align*}
\end{theorem}
The next result  the loss in performance of a policy that is greedy w.r.t. $J_{\tu}$:
\begin{theorem}[Theorem~$3.1$ of \cite{ALP}]
\label{restatepol}
Let $\tilde{u}$ be the greedy policy with respect to the solution $\tj$ of the ALP. Then,
\begin{align*}
||J^*-J_{\tu}||_{1,c}\leq \frac{1}{1-\alpha}||J^*-\tj||_{1,c'}\,,
\end{align*}
where $c'=(1-\alpha)c^\top(I-\alpha P_{\tu})^{-1}$.
\end{theorem}
Theorems~\ref{restateval} and ~\ref{restatepol} together imply that the ALP addresses both the control and prediction problems. Please refer to \cite{ALP} for a more detailed treatment of the ALP.\\
Note that the ALP is a linear program in $k$ ($\ll n$) variables as opposed to the LP in \eqref{mdplpshort} which has $n$ variables. Nevertheless, the ALP has $nd$ constraints (same as the LP) which is an issue when $n$ is large and calls for constraint approximation/reduction techniques.
\todoc{Column generation, Dantzig-Wolf decomposition?}
\end{comment}

### 2016-07-08 mdp.tex
Chandru's shortened version

%!TEX root =  autocontgrlp.tex
\section{Markov Decision Processes (MDPs)}
In this section, we briefly discuss the basics of Markov Decision Processes (MDPs) (the reader is referred to \cite{BertB,Puter} for a detailed treatment).\\
An MDP is a $4$-tuple $\langle S,A,P,g\rangle$, where $S$ is the state space, $A$ is the action space, $P$ is the probability transition kernel and $g$ is the reward function. We consider MDPs with large but finite number of states. We also assume that the number of actions is finite.
Without the loss of generality, we set  $S=\{1,2,\ldots,n\}$  and the action set is given by $A=\{1,2,\ldots,d\}$. For simplicity, we assume that all actions are feasible in all states. 
The probability transition kernel $P= (p_a)_{a\in A}$ specifies the probability $p_a(s,s')$ of transitioning from state $s$ to state $s'$ under the action $a$. We denote the reward obtained for performing action $a\in A$ in state $s\in S$ by $g_a(s)$.
By a policy, we mean a sequence $\pi=\{\pi_0,\ldots,\pi_n,\ldots\}$ of functions $\pi_n, n\geq 0$ 
	that describe the manner in which an action is picked in a given state at time $n$. 
In the most general case, $\pi_n$ maps $(S \times A)^n \times S$, i.e., 
	histories $h_n =(s_0,a_0,\dots,s_{n-1},a_{n-1},s_n)$ 
	of state-action sequences of length $2n+1$, to a distribution over the actions. 
We denote the class of policies by $\Pi$. 
The infinite horizon discounted reward corresponding to state $s$ under a policy $\pi$ is denoted by $J_\pi(s)$ and is defined by \todoc{Explain the meaning of $\eqdef$.}
\begin{align}
J_\pi(s) \eqdef
	\E\left[\sum_{n=0}^\infty \alpha^n g_{a_n}(s_n)\,\Large|\,\pi\right]\,,\nn
\end{align}
	where $a_{n} \sim \pi_n(\,\cdot\,|s_0,a_0,\dots,s_{n-1},a_{n-1},s_n)$, $s_{n+1} \sim p_{a_n}(s_n,\cdot)$ 
	and $\alpha \in (0,1)$ is a given discount factor. 
Here $J_\pi(s)$ is known as the value of the state $s$ under the policy $\pi$, and $J_\pi:S\to \R$ is 
 %the vector quantity $J_\pi\eqdef(J_\pi(s))_{s\in S}\in R^n$ is 
 called the value-function corresponding to the policy $\pi$. 
 The optimal value function $J^*:S\to \R$ is defined as $J^*(s) \eqdef
 \sup_{\pi\in\Pi} J_\pi(s)$, $s\in S$.
 %\underset{\pi \in \Pi}{\sup}\, J_\pi(s)$, $s\in S$.

A stationary deterministic policy (SDP) is one where $\pi_n\equiv u$ for all $n\geq 0$ for some $u\colon S \ra A$
in the sense that $\pi_n(a|s_0,a_0,\dots,s_{n-1},a_{n-1},s_n)$ is one for $a = u(s_n)$ and is zero elsewhere.
By abuse of notation we denote such an SDP by $u$ itself instead of $\pi$. 
In the setting that we consider, one can find an SDP that is optimal \cite{BertB,Puter}, i.e., there exists an SDP $u^*$ such that $J_{u^*}(s)=J^*(s),\forall s\in S$. 
In this paper, we restrict our focus to the class $U$ of SDPs. 
%	and without loss of generality 
%	we assume that there exists a unique SDP $u^*$ that is optimal.  
	\todoc{Do we need uniqueness? I got rid of it.}
%Under a stationary policy $u$ (or $\pi$), 
%	the MDP is a Markov chain and we denote its probability transition kernel 
%	by $P_u=(p_{u(s)}(s,s'),s,s'=1,\ldots,n)$ (or $P_\pi=(p_{\pi(s)}(s,s'),s,s'=1,\ldots,n)$, 
%	where $p_{\pi(s)}(s,s')=\sum_{a\in A}\pi(s,a)p_a(s,s')$ and $\pi(s)=(\pi(s,a), a\in A)$).
Under a stationary deterministic policy $u$, 
	the MDP gives rise to a Markov chain with state space $S$,
whose  transition probability kernel we will denote
	by $P_u=(p_{u(s)}(s,s'),s,s'=1,\ldots,n)$.  \todoc{I don't think we need stationary \emph{stochastic} policies anywhere.}

Given an MDP, our aim is to find the optimal value function $J^*$ and an optimal policy SDP $u^*$. 
Any optimal policy $u^*$ and the value function $J^*$ obey the Bellman equation (BE): for all $ s \in S$, 
\begin{subequations}\label{bell}
\begin{align}
\label{bellval}J^*(s)&=\max_{ a\in A}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')J^*(s')\big),\\
\label{bellpol}u^*(s)&\in \argmax_{ a\in A}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')J^*(s')\big)\,.
\end{align}
\end{subequations}
If $J^*$ is computed first, then $u^*$ can be obtained via \eqref{bellpol}.

For convenience, in what follows we identify $\R^n$ with $\R^{S}$:
Any vector $J\in \R^n$ is viewed as a real-valued function over $S$ via $J(s) = J_s$ and vice versa.
We now define some operators acting on $\R^n$ (equivalently, on $\R^S$) that will be useful later.
The Bellman operator $T$ is defined using the model parameters of the MDP as follows:
\begin{definition}
The Bellman operator $T: \R^n \to \R^n$ is defined by 
\begin{align}
(TJ)(s)=\max_{a \in A}\big(g_a(s)+\alpha \sum_{s'} p_a(s,s')J(s')\big)\,,  \qquad s\in S\,,J\in \R^n\,.
\end{align}
\end{definition}
Similarly one can define the Bellman operator restricted to an SDP $u$ as follows:
\begin{definition}
The Bellman operator $T_u:\R^n \to \R^n$  restricted to an SDP $u$ is defined by
\begin{align}
(T_uJ)(s)=g_{u(s)}(s)+\alpha \sum_{s'} p_{u(s)}(s,s')J(s')\,,  \qquad s\in S\,,J\in \R^n\,.
\end{align}
\end{definition}
Given $J \in \R^n$, $TJ$ is the `one-step' greedy value function. It is also useful to define the notion of a one-step greedy policy as below:
\begin{definition}
A policy $\tilde{u}$ is said to be greedy with respect to $\tilde{J}$ if
\begin{align}\label{subpol}
\tilde{u}(s)\in \underset{a \in A}{\argmax} \, \big(g_a(s)+\alpha\sum_{s'} p_a(s,s')\tilde{J}(s')\big)\,.
\end{align}
\end{definition}
We also define the Bellman operator $H$ for action values  \cite{BertB}: 
\begin{definition}\label{bellactval}
Let $H: \R^n \to \R^{nd}$ be defined as follows: For $J\in \R^n$,
\begin{align}
HJ&=\left[ {\begin{array}{c} H_1 J  \\ \vdots \\ H_d J\end{array}} \right]\in \R^{nd}, \text{ where}\nn\\
(H_a J)(s)&= g_a(s)+\alpha \sum_{s'}p_a(s,s') J(s'), \qquad s\in S, a\in A.
\end{align}
\end{definition}
We now state without proof the most important properties of the Bellman operator. 
Though in these results, we make use of the Bellman operator $T$, the results also trivially hold for $T_u$ as well.

%\subsection{Properties of $T$}
\begin{lemma}\label{maxnorm}
$T$ is a $\max$-norm contraction operator with contraction factor $\alpha$, i.e., given $J_1, J_2 \in \mathbf{R}^n$,
\todoc{Introduce $\norm{\cdot}_{\infty}$. Should we have a notation/symbols section?}
\begin{align}
||TJ_1-TJ_2||_\infty\leq \alpha ||J_1-J_2||_\infty.
\end{align}
\end{lemma}
\begin{lemma}\label{uniquesol} 
$J^*$ is a unique fixed point of $T$, i.e., $J^*=TJ^*$.
\end{lemma}
For $J_1,J_2\in \R^n$, we write $J_1\le J_2$ if $J_1(s)\le J_2(s)$ holds for all $s\in S$.
The Bellman operator $T$ exhibits two more important properties presented in the following lemmas (see \cite{BertB} for proofs):
\begin{lemma}\label{monotone}
$T$ is a monotone map, i.e., given $J_1,J_2 \in \R^n$ such that $J_1\leq J_2$, we have $T J_1\leq T J_2$. 
Further, if $J\in \R^n$ is such that $J\geq TJ$ then $J\geq J^*$.
\end{lemma}
\begin{lemma}\label{shift}
Given $J\in \R^n$, $t \in \R$ and $\mathbf{1} \in \R^n$, a vector with all entries $1$, we have
\begin{align}
T(J+t\mathbf{1})=TJ+\alpha t\mathbf{1}.
\end{align}
\end{lemma}
Note that Lemmas~\ref{maxnorm}--\ref{shift} also hold for the Bellman operator $H$ defined for the action values in Definition~\ref{bellactval}.\par
Solving an MDP involves handling two sub-problems namely the problem of \emph{control} and the problem of \emph{prediction}. The problem of \emph{control} deals with coming up with a good (and if possible an optimal) policy. Often, in order to solve the problem of \emph{control}, one needs to solve the problem of \emph{prediction}, which deals with computing the value function $J_u$ of the policy $u$. The fact that the two problems are related is reflected in the Bellman equation in \eqref{bell}, where $J^*$ from \eqref{bellval} is used in \eqref{bellpol} to obtain an optimal policy $u^*$. Thus, the Bellman equation is at the heart of the solution methods to MDPs. Any solution method to MDP is said to be complete only if it satisfactorily (with provable performance guarantees) addresses both the prediction and the control problems. 
\todoc{Why insist on prediction?}

The basic solution methods namely value iteration, policy iteration and linear programming (LP) \cite{BertB} solve both the control and prediction problems. Of the three basic methods, in this paper, we focus on the LP formulation given by
\begin{align}\label{mdplp}
\begin{split}
\min_{J\in \R^n}\, &c^\top J\\
\text{s.t.}\mb &J(s)\geq g_a(s)+\alpha\sum_{s'}p_a(s,s')J(s'), \qquad s\in S, a \in A,
\end{split}
\end{align}
where $c\in \R^n$ is any vector whose components are all positive.
\todoc{I changed this from nonnegative to positive.} 
One can show that $J^*$ is the solution to the linear program given by \eqref{mdplp} \cite{BertB}. 
Also, of the three methods, value iteration and LP formulation are value function based methods, i.e., they compute $J^*$ directly and then $u^*$ is obtained by plugging $J^*$ in \eqref{bellpol}.

While the basic methods (i.e., VI, PI and LP) can be used to compute exact values $J^*$ and $u^*$ for MDPs with a small number of states, they are intractable in the case of MDPs with a large number of states.

>>>>

%!TEX root =  autocontgrlp.tex
\section{Markov Decision Processes (MDPs)}
We consider MDPs with large but finite number of states. Thus $S=\{1,2,\ldots,n\}$ for some large $n$ is the state space, and the action set is given by $A=\{1,2,\ldots,d\}$. For simplicity, we assume that all actions are feasible in all states. The probability transition kernel $P$ specifies the probability $p_a(s,s')$ of transitioning from state $s$ to state $s'$ under the action $a$. We denote the reward obtained for performing action $a\in A$ in state $s\in S$ by $g_a(s)$.\par
A stationary deterministic policy\footnote{For the scope of this paper, it suffices to restrict our attention to stationary deterministic policies.}(SDP) or simply a policy is a map $u\colon S\ra A$ which specifies the action selection mechanism. Given an SDP $u$, the infinite horizon discounted reward corresponding to state $s$ under $u$ is denoted by $J_u(s)$ and is defined by
\begin{align}
J_u(s)\stackrel{\Delta}{=}\E[\sum_{n=0}^\infty \alpha^n g_{a_n}(s_n)|s_0=s,a_n=u(s_n)\mbox{ }\forall n\geq 0],\nn
\end{align}
where $\alpha \in (0,1)$ is a given discount factor. Here $J_u(s)$ is known as the value of the state $s$ under SDP $u$, and the vector quantity $J_u\eqdef(J_u(s), \forall s\in S)\in R^n$ is called the value function corresponding to the SDP $u$. The \emph{optimal policy} $u^*$ is obtained as $u^*(s)\eqdef\arg\max_{u\in U}J_u(s)$\footnote{An optimal (SDP) $u^*$ exists and is well defined in the case of infinite horizon discounted reward MDP, for more details see \cite{BertB}.}, where $U$ is the class of all SDPs. The \emph{optimal value function} $J^*$ is the one obtained under the optimal policy, i.e., $J^*=J_{u^*}$.\par
Any optimal policy $u^*$ and value function $J^*$ obey the Bellman equation (BE): for all $ s \in S$, 
\begin{subequations}\label{bell}
\begin{align}
\label{bellval}J^*(s)&=\max_{ a\in A}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')J^*(s')\big),\\
\label{bellpol}u^*(s)&= \underset{a\in A}{\argmax}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')J^*(s')\big)\,.
\end{align}
\end{subequations}
Once $J^*$ is computed, $u^*$ can be obtained via \eqref{bellpol}. Thus computing the value function is at the heart of most solution methods to MDP.\par
For convenience, in what follows we allow any vector $J\in \R^n$ to be viewed as a real-valued function over $S$ via $J(s) = J_s$ and vice versa.
We now define some operators acting on $\R^n$ (equivalently on $\R^S$) that will be useful later.
%The Bellman operator $T$ is defined using the model parameters of the MDP as follows:
\begin{definition}
Let $c,\rho,\chi:S \to \R_+$ be positive valued functions, where $\R_+$ denotes the set of strictly positive reals. Then for $J\in \R^n$, $a\in A$ and $s\in S$, 
define
\begin{enumerate}[(i)]
\item The Bellman operator $T\colon \R^n \ra \R^n$ as $(TJ)(s)=\max_{a \in A}\big(g_a(s)+\alpha \sum_{s'} p_a(s,s')J(s')\big).
$
\item \label{bellactval} The Bellman operator (of action values) $H: \R^n \to \R^{nd}$ for state-action values as $HJ=[ H_1 J,\cdots,H_d J]^\top\in \R^{nd},$ where $(H_a J)(s)= g_a(s)+\alpha \sum_{s'}p_a(s,s') J(s')$.
\item The weighted $L_1$-norms $\norm{\cdot}_{1,c}$ 
and 
the weighted $L_\infty$-norms  $\norm{\cdot}_{\infty,\rho}$ as $
||J||_{1,c}=\sum_{s \in S} c(s)|J(s)|, 
||J||_{\infty,\rho}=\max_{s \in S} \frac{|J(s)|}{\rho(s)}$.
\item The discounted maximal inflation of $\chi$ due to $P = (p_a)_{a\in A}$ as $\beta_{\chi}=\max_{s \in S} \frac{\underset{a \in A}{\max}\big(\alpha\sum_{s'}p_a(s,s')\chi(s')\big)}{\chi(s)}$.
\item The function $\chi:S\to\R_+$ above to be a \emph{Lyapunov} function for $P = (p_a)_{a\in A}$ if $\beta_{\chi}<1$.
\item $E$ to be the $nd\times n$ matrix given by $E=[I,\ldots,I]^\top$, i.e., $E$ is obtained by stacking $d$ identical $n\times n$ identity matrices one over the other.
\item The greedy policy with respect to the value function $\tj$ as \begin{align*}\tu\eqdef\underset{a\in A}{\argmax}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')\tj(s')\big).\end{align*}
\end{enumerate}
\end{definition}
We now state without proof the most important properties of the Bellman operator.
The proofs are immediate from the definitions, but can also be found in \cite{BertB}.
First, we introduce some extra notation:
For $J_1,J_2\in \R^n$, we write $J_1\le J_2$ if $J_1(s)\le J_2(s)$ holds for all $s\in S$.
We use $\one \in \R^n$ to denote a vector with all entries $1$.
The maximum norm $\norm{\cdot}_{\infty}$ is defined by $ \norm{v}_{\infty} = \max_{s\in S} |v(s)|$.
\begin{lemma}\label{monotone}
$T$ is a monotone map, i.e., given $J_1,J_2 \in \R^n$ such that $J_1\leq J_2$, we have $T J_1\leq T J_2$. 
\end{lemma}
\begin{lemma}\label{shift}
Given $J\in \R^n$ and $t \in \R$, we have
\begin{align}\label{eq:shift}
T(J+t\one)=TJ+\alpha t\one.
\end{align}
\end{lemma}
\begin{lemma}\label{maxnorm}
If $T: \R^n \to \R^n$ is any operator that is monotonous and satisfies~\eqref{eq:shift} then 
$T$ is a $\max$-norm contraction operator with contraction factor $\alpha \in (0,1)$, i.e., given $J_1, J_2 \in \mathbf{R}^n$,
\begin{align}
||TJ_1-TJ_2||_\infty\leq \alpha ||J_1-J_2||_\infty.
\end{align}
\end{lemma}
\begin{lemma}\label{uniquesol} 
$J^*$ is a unique fixed point of $T$, i.e., $J^*=TJ^*$.
\end{lemma}
\begin{corollary}
If $J\in \R^n$ is such that $J\geq TJ$ then $J\geq TJ^2\geq \ldots \geq J^*$.
\end{corollary}
Though in \cref{monotone,shift,maxnorm}, we make use of the Bellman operator $T$, the results also hold for $H$ as well.
\subsection{Linear Programming Formulation}
The optimal value function $J^*$ can be obtained by solving the following linear program:
\begin{align}\label{mdplp}
\min_{J\in \R^n}\, &c^\top J\\
\text{s.t.}\mb &J(s)\geq g_a(s)+\alpha\sum_{s'}p_a(s,s')J(s'),\nn\\
&\forall s \in S, a \in A,
\end{align}
where $c$ is a probability distribution such that $c(s)>0, \forall s \in S$ and $\sum_{s\in S}c(s)=1$.\par
When the MDP has a large number of states, it is difficult to solve for $J^*$ using either the linear program \eqref{mdplp} or other full state representation methods such as value iteration or policy iteration \cite{BertB}. A practical solution is to resort to function approximation. The most widely used is the linear function approximation wherein the exact value function is approximated by a function that is obtained as a linear combination of a set of pre-selected basis functions.
